---
title: "Plus A"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "50%")
library(resampledata)
```

##Plus A: Earthquake time difference case study

###Preparation: Defining some helper functions

We must first define some functions to help us with the calculations of the parameter estimates.

We define the function <tt>weibull.shape</tt> which takes input shape parameter k and
the data to compute
\[
f(k, \mbox{data}) = 
(1/k)+ (1/n)*\sum (\log(x_i)) +(1/\alpha)\sum x_i^k\log(x_i)
\]
where $\alpha= \sum x_i^k$.

```{r}
weibull.shape <- function(k, data)
{
  numer <- colSums(outer(data, k, "^") * log(data))
  denom <- colSums(outer(data, k, "^"))
  numer/denom - 1/k - mean(log(data))
}
```

####Another helper function
This function takes input shape parameter k
 and data to compute
  $k^{th}$ root of $(1/n) \sum x_i^k$, where 
 n=number of data values.

```{r}
weibull.scale <- function(k, data)
{
  mean(data^k)^(1/k)
}
```

####Built-in helper function
`uniroot` is a built-in R function which estimates the root of a function.

User provide this function any arguments needed for function, and a guess of two values a, b around root.
If `lower = a, upper = b` are your two guesses,
f(a) and f(b) must be opposite signs.

###Now, the analysis

The data set <tt>Turbine</tt> is in the <tt>resampledata</tt> package (loaded in set-up step above.)

We use the <tt>uniroot</tt> command to solve for k:

\[
f(k, \mbox{data}) = 
(1/k)+ (1/n)*\sum (\log(x_i)) +(1/\alpha)\sum x_i^k\log(x_i) = 0
\]

```{r}
time <- Quakes$TimeDiff
#alternatively, wind <- subset(Turbine, select=AveSpeed, drop=TRUE)

#estimate the shape parameter k
k.est <- uniroot(weibull.shape, data = time, lower = 0.8,upper = 1)$root
```
The output $root gives use the estimate of k (`r k.est`).

With estimate of shape parameter k, now find estimate
of scale parameters $\lambda$.

```{r}
lamda.est <- weibull.scale(k.est, time)
```

Now we have esimates for k and $\lambda$ (`r lamda.est`).

Plot histogram with density curve overlap
The <tt>prob=TRUE</tt> argument scales histogram to area 1.

```{r, fig.align='center'}
hist(time, main = "Distribution of average time difference between earthquakes",
     xlab = "days", prob = TRUE, ylim=c(0, 0.06))

curve(dweibull(x, k.est, lamda.est), add = TRUE, col = "blue", lwd = 2)
```

For the empirical cdf:

```{r, fig.align='center'}
plot.ecdf(time,main = "ECDF of time difference")
curve(pweibull(x,k.est,lamda.est), add = TRUE, col = "blue", lwd = 2)
```
